Diagnosing Child Poverty Reduction Interventions

### 2024-09-20 minos/outcomes/CPR_Intervention_Diagnostics.Rmd
Split scaling implemented; scaling is now lopsided with reduced variance at the bottom of the distribution

```{r}
# Categorize SF-12 MCS transitions: "Improving", "Stable", "Worsening"
prepare_sf12_transitions <- function(data) {
  data %>%
    select(pidp, time, income_quint, SF12_MCS) %>%
    group_by(pidp) %>%
    arrange(time) %>%  # Ensure data is ordered by time
    mutate(SF12_MCS_change = SF12_MCS - lag(SF12_MCS)) %>%
    mutate(SF12_MCS_transition = case_when(
      SF12_MCS_change > 0 ~ "Improving",
      SF12_MCS_change == 0 ~ "Stable",
      SF12_MCS_change < 0 ~ "Worsening",
      TRUE ~ NA_character_
    )) %>%
    filter(!is.na(SF12_MCS_transition)) %>%
    ungroup()
}

# Prepare the data
sf12_transitions <- prepare_sf12_transitions(data)

# Calculate the count of transitions per quintile and transition type
sf12_transition_counts <- sf12_transitions %>%
  group_by(income_quint, SF12_MCS_transition) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  spread(SF12_MCS_transition, count, fill = 0)

# Normalize the counts to show proportions per quintile
sf12_transition_counts <- sf12_transition_counts %>%
  gather(SF12_MCS_transition, count, -income_quint) %>%
  group_by(income_quint) %>%
  mutate(proportion = count / sum(count)) %>%
  ungroup()

# Plot the heatmap
ggplot(sf12_transition_counts, aes(x = SF12_MCS_transition, y = as.factor(income_quint), fill = proportion)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "red") +
  labs(title = "SF-12 MCS Transitions Heatmap by Quintile",
       x = "SF-12 MCS Transition", y = "Income Quintile", fill = "Proportion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

# Example usage
plots <- plot_sf12_heatmap_difference(base.dat, int.dat)
plots$count_heatmap  # Display the count heatmap
plots$prop_heatmap  # Display the proportion heatmap

>>>>

```{r}
# Categorize SF-12 MCS transitions: "Improving", "Stable", "Worsening"
# prepare_sf12_transitions <- function(data) {
#   data %>%
#     select(pidp, time, income_quint, SF12_MCS) %>%
#     group_by(pidp) %>%
#     arrange(time) %>%  # Ensure data is ordered by time
#     mutate(SF12_MCS_change = SF12_MCS - lag(SF12_MCS)) %>%
#     mutate(SF12_MCS_transition = case_when(
#       SF12_MCS_change > 0 ~ "Improving",
#       SF12_MCS_change == 0 ~ "Stable",
#       SF12_MCS_change < 0 ~ "Worsening",
#       TRUE ~ NA_character_
#     )) %>%
#     filter(!is.na(SF12_MCS_transition)) %>%
#     ungroup()
# }
# 
# # Prepare the data
# sf12_transitions <- prepare_sf12_transitions(data)
# 
# # Calculate the count of transitions per quintile and transition type
# sf12_transition_counts <- sf12_transitions %>%
#   group_by(income_quint, SF12_MCS_transition) %>%
#   summarise(count = n()) %>%
#   ungroup() %>%
#   spread(SF12_MCS_transition, count, fill = 0)
# 
# # Normalize the counts to show proportions per quintile
# sf12_transition_counts <- sf12_transition_counts %>%
#   gather(SF12_MCS_transition, count, -income_quint) %>%
#   group_by(income_quint) %>%
#   mutate(proportion = count / sum(count)) %>%
#   ungroup()
# 
# # Plot the heatmap
# ggplot(sf12_transition_counts, aes(x = SF12_MCS_transition, y = as.factor(income_quint), fill = proportion)) +
#   geom_tile() +
#   scale_fill_gradient(low = "white", high = "red") +
#   labs(title = "SF-12 MCS Transitions Heatmap by Quintile",
#        x = "SF-12 MCS Transition", y = "Income Quintile", fill = "Proportion") +
#   theme_minimal() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

# Example usage
#plots <- plot_sf12_heatmap_difference(base.dat, int.dat)
#plots$count_heatmap  # Display the count heatmap
#plots$prop_heatmap  # Display the proportion heatmap
